---
title: "Tidyverse Exercise"
author: "Claudi Lleyda Moltó"
date: "1462908"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "")

library(tidyverse)
library(nycflights13)

flights.tib <- as_tibble(flights)
```

 (1) Using the `flights` dataset, find all flights that

  - Had an arrival delay of two or more hours

```{r}
flights.tib %>% filter(arr_delay >= 120) %>% select(flight, arr_delay)
```

  - Flew to Houston

```{r}
houston <- c("IAH", "HOU")
flights.tib %>% filter(dest %in% houston) %>% select(flight, dest)
```

  - Arrived more than two hours late, but didn't leave late

```{r}
flights.tib %>% filter(arr_delay > 120, dep_delay <= 0) %>%
                select(flight, dep_delay, arr_delay)
```

  - Were delayed by at least an hour, but made up over $30$ minutes in flight

```{r}
flights.tib %>% filter(dep_delay >= 60, dep_delay - arr_delay > 30) %>%
                select(flight, dep_delay, arr_delay)
```

 2. Another useful dplyr filtering helper is `between()`. What does it do? Can you use it to simplify the code needed to answer the previous challenges?

The function `between(x, left, right)` as a shortcut for `x >= left & x <= right`. The documentation for the function promises it is an efficient implementation that is compatible with `SQL` queries for remote tables, which could be a huge benefit in such contexts.

In this case the code is longer and more verbose, and is cumbersome to use with non-bounded  or non-closed intervals (since it does not allow the use of non-strict inequalities like `x > left`, but we can translate them to `x >= left+1` in the case of `int` values), but there is a chance it will run faster.

```{r}
# filter(arr_delay >= 120)
flights.tib %>% filter(between(arr_delay, 120, Inf)) %>% select(flight, arr_delay)
# filter(arr_delay > 120, dep_delay <= 0)
flights.tib %>% filter(
  between(arr_delay, 120+1, Inf),
  between(dep_delay, -Inf, 0)
) %>% select(flight, arr_delay, dep_delay)
# filter(dep_delay >= 60, dep_delay - arr_delay > 30)
flights.tib %>% filter(
  between(dep_delay, 60, Inf),
  between(dep_delay - arr_delay, 30+1, Inf)
) %>% select(flight, arr_delay, dep_delay)
```

 3. Sort flights to find the fastest flights.

```{r}
flights.tib %>% mutate(speed = distance / air_time) %>% arrange(-speed) %>%
                select(flight, distance, air_time, speed)
```

 4. Create a new data frame having variables with the `dep` string.

```{r}
flights.tib %>% select(contains("dep")) %>% names()
```

 5. Create a new data frame having the hour and minute of departure

```{r}
flights.tib %>% mutate(
  dep_hour = as.integer(dep_time %/% 100),
  dep_minute = as.integer(dep_time %% 100)
) %>% select(flight, dep_time, dep_hour, dep_minute)
```
