---
title: "Markdown Exercise"
author: "Claudi Lleyda Moltó"
date: "1462908"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment="")

library(tidyverse)
library(ggplot2)
library(datasets)

mtcars$vs <- as.factor(mtcars$vs)
levels(mtcars$vs)=c("V-shaped","Straight")
mtcars$am <- as.factor(mtcars$am)
levels(mtcars$am)=c("Automatic", "Manual")
```

The `mtcars` dataset was extracted from the $1974$ Motor Trend US magazine. It contains data from $32$ different automobiles to compare $10$ aspects of automobiles, like fuel consumption. All the automobiles were from $1973$-$1974$.

We will work with the following variables:

 - `mpg`: Miles per gallon (`num`): Fuel consumption
 - `wt`: Weight (`num`): Weight of the automobile in $1000\textrm{lb}$ scale ($453.59\textrm{kg}$)
 - `vs` Engine (`cat`): Either V-shaped or straight
* `am`: Transmission (`cat`): Either automatic or manual
* `hp`: Gross horsepower (`num`): Gross horsepower
* `qsec`: 1/4 mile time (`num`): Time to travel a quarter mile

What follows is some entries of the dataset:

```{r}
tail(mtcars)
```

If we tabulate the relation between the two categorical variables (`vs` and `am`),

```{r}
table(mtcars$vs, mtcars$am)
```

we observe that they could be correlated, but we have too few samples.

To have an idea of the distribution of the continuous variables (`mpg`, `wt`, `hp` and `qsec`), we plot their histograms.

```{r}
par(mfrow = c (2,2))
hist(mtcars$mpg, xlab = "Miles/gallon", main = "Histogram of Miles/gallon")
hist(mtcars$wt, xlab = "Weight", main = "Histogram of Weight")
hist(mtcars$hp, xlab = "Gross horsepower", main = "Histogram of Gross horsepower")
hist(mtcars$qsec, xlab = "1/4 miles time (s)", main = "Histogram of 1/4 miles time")
```

Finally, we want to see if a linear regression is a good model to relation fuel consumption (`mpg`) and gross horsepower (`hp`).

```{r}
ggplot(mtcars, aes(x = mpg, y = hp)) +
  geom_point() +
  stat_smooth(formula = y ~ x, method = "lm", col = "red")
```

Again, there are too few samples to define properly the linear regression and to test if it works quite well, but it's seems that could be a good option.

